---
title: "Exploring the UDBP Variable"
author: "Windy Wang"
date: '`r Sys.Date()`'
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, echo = FALSE, message = FALSE, warning = FALSE)
library(dplyr)
devtools::load_all()
load_data(update = TRUE)
set_options()
# extrafont::loadfonts(device="win")

ds <- project_data
```

## Descriptive Statistics

Below shows the distribution of urinary vitamin D binding protein
```{r histogram}

ds %>% 
  histo_plot(variable = "UDBP", 
             bin = 30, 
             xlab = "Urinary VDBP (ng/mL)",
             facet = TRUE)
```

Summary statistics at baseline
```{r summary baseline}

ds %>% 
  filter(fVN == "Baseline") %>% 
  select(UDBP) %>% 
  summary()
```

Summary statistics at 3 year
```{r summary 3year}

ds %>% 
  filter(fVN == "3Year") %>% 
  select(UDBP) %>% 
  summary()
```

Summary statistics at 6 year
```{r summary 6year}
ds %>% 
  filter(fVN == "6Year") %>% 
  select(UDBP) %>% 
  summary()
```

Boxplot of log(UDBP:cr) change over time
```{r boxplot}

ds %>% 
  box_plot(xvar="fVN",
           yvar="log(udbpCrRatio)",
           xlab="Visit Number",
           ylab="log Urinary VDBP (ng/mL)")
```


## UDBP Progression

```{r long to wide}

# Convert UDBP long-data to wide

ds_wide_udbp <- ds %>% 
  dplyr::select(SID, VN, UDBP) %>% 
  tidyr::spread(key = VN, value = UDBP, drop = TRUE, sep = ".") %>% 
  dplyr::rename(udbp1 = VN.1,
                udbp3 = VN.3,
                udbp6 = VN.6) %>% 
  dplyr::mutate(
    udbp_change = ifelse(udbp3 > udbp1 & udbp6 > udbp1, "progress", "regress"))

```

Growth curve for everyone
```{r growth curve}
ds %>% 
  dplyr::mutate(fSID = factor(SID)) %>%
  line_plot(xvar="fVN",
            yvar="UDBP",
            byvar="fSID",
            xlab="Visit Number",
            ylab="Urinary VDBP (ng/mL)")
```

Zoomed in for UDBP < 25 ng/mL
```{r growth curve zoom}
ds %>% 
  dplyr::filter(UDBP < 25) %>% 
  dplyr::mutate(fSID = factor(SID)) %>%
  line_plot(xvar="fVN",
            yvar="UDBP",
            byvar="fSID",
            xlab="Visit Number",
            ylab="Urinary VDBP (ng/mL)")
```

UDBP change with ACR
```{r growth curve acr}
ds %>% 
  dplyr::mutate(fSID = factor(SID)) %>%
  line_plot(xvar="acrStatus",
            yvar="UDBP",
            byvar="fSID",
            xlab="ACR",
            ylab="Urinary VDBP (ng/mL)")
```

UDBP change with eGFR
```{r growth curve eGFR}
ds %>% 
  dplyr::mutate(fSID = factor(SID)) %>%
  line_plot(xvar="eGFRStatus",
            yvar="UDBP",
            byvar="fSID",
            xlab="eGFR",
            ylab="Urinary VDBP (ng/mL)")
```

Number of progressors (UDBP value was higher at subsequent visits compared to baseline)
*NOTE: having trouble with missingness in coding
```{r progressors}

ds_wide_udbp %>% 
  dplyr::group_by(udbp_change) %>% 
  dplyr::summarise(n = n())
```


<!-- Mixed models uses long-format data -->

<!-- ```{r latent curve mixed models} -->

<!-- # Random slope model - assumes trajectory is different from person-to-person -->
<!-- m <- lme4::lmer(UDBP ~ (VN | SID), data=ds) -->
<!-- summary(m) -->

<!-- # Confidence intervals -->
<!-- dat.new <- data.frame(time=1:3) -->
<!-- dat.new$measure <- predict(m, dat.new, re.form=NA) -->
<!-- m.mat <- model.matrix(terms(m), dat.new) -->
<!-- dat.new$var <- diag(m.mat %*% vcov(m) %*% t(m.mat)) + VarCorr(m)$subject.id[1] -->
<!-- dat.new$pvar <- dat.new$var + sigma(m)^2 -->
<!-- dat.new$ci.lb <- with(dat.new, measure - 1.96*sqrt(var)) -->
<!-- dat.new$ci.ub <- with(dat.new, measure + 1.96*sqrt(var)) -->
<!-- dat.new$pi.lb <- with(dat.new, measure - 1.96*sqrt(pvar)) -->
<!-- dat.new$pi.ub <- with(dat.new, measure + 1.96*sqrt(pvar)) -->

<!-- # Plot -->
<!-- ds %>%  -->
<!--   ggplot2::ggplot(ggplot2::aes(x=time, y=measure)) +  -->
<!--   geom_line(data=dat, alpha=.02, aes(group=subject.id)) +  -->
<!--   geom_errorbar(width=.02, colour='red',  -->
<!--                 aes(x=time-.02, ymax=ci.ub, ymin=ci.lb)) + -->
<!--   geom_line(colour='red', linetype='dashed', aes(x=time-.02)) +  -->
<!--   geom_point(size=3.5, colour='red', fill='white', aes(x=time-.02)) +    -->
<!--   geom_errorbar(width=.02, colour='blue',  -->
<!--                 aes(x=time+.02, ymax=pi.ub, ymin=pi.lb)) + -->
<!--   geom_line(colour='blue', linetype='dashed', aes(x=time+.02)) +  -->
<!--   geom_point(size=3.5, colour='blue', fill='white', aes(x=time+.02)) +  -->
<!--   theme_bw() -->
<!-- ``` -->

