---
title: "Paper 1: Kidney Dysfunction"
author: "Windy Wang"
output: word_document
---

```{r setup, collapse=TRUE, echo=FALSE, message=FALSE, warning=FALSE}

# Only set if the Rmd file is not in the parent directory (ie. 'projectname/')
knitr::opts_knit$set(root.dir = '../')

knitr::opts_chunk$set(collapse = TRUE, echo = FALSE, message = FALSE, warning = FALSE)
library(tidyverse)
library(PROMISE.data)
devtools::load_all()
# load_data(update = TRUE)
set_options()
# extrafont::loadfonts(device="win")
```

```{r loadData, message=FALSE}

# source('.Rprofile')
# run_setup()
# load_data('data/project_data.rda')
# dim(ds)
ds <- project_data

dsBase <- ds %>% 
  dplyr::filter(fVN == "Baseline")

# Subjects with measurements at all visit numbers

dsComplete <- ds %>%
  dplyr::group_by(SID) %>%
  dplyr::filter(n() == 3) %>%
  dplyr::ungroup()

```

```{r caption_setup}

## Include captions below using `captioner` package

fig_nums <- captioner::captioner(prefix = 'FIGURE')
supfig_nums <- captioner::captioner(prefix = 'SUPPLEMENTARY FIGURE')
tab_nums <- captioner::captioner(prefix = 'TABLE')
suptab_nums <- captioner::captioner(prefix = 'SUPPLEMENTARY TABLE')
```

PAPER 1: KIDNEY DYSFUNCTION RESULTS
===================================

### Subject Characteristics

```{r subchar udbp tertile}

# tab_nums(kidney_subchar_baseline, "Subject characteristics across uVDBP tertiles at baseline")

dsBaseNoBPMeds <- dsBase %>% 
  dplyr::filter(fMedsBP == "No")

subchar_table_tert <- tableone::CreateTableOne(
  vars = c("Age",
           "Sex",
           "Ethnicity",
           "BMI",
           "Waist",
           "eGFR",
           "ACR",
           "UrineCreatinine",
           "UrineMicroalbumin",
           "UrinaryCalcium",
           "UDBP",
           "udbpCr",
           "Systolic",
           "Diastolic",
           "MeanArtPressure",
           "dmStatus"),
  strata = c("udbpTertile"),
  data = dsBaseNoBPMeds,
  factorVars = c("Sex", "Ethnicity", "dmStatus")
) %>% 
  print(nonnormal = c("UDBP",
                      "ACR",
                      "UrineMicroalbumin"),
        quote = FALSE,
        noSpaces = TRUE)

# explore non-linear variables (delete later)

dsBase %>% 
  scatter_plot("udbpCr", "BMI",
               "uVDBP:cr", "BMI")

# explore (delete later)

dsBase %>% 
  summarise(n = n())

```

TABLE 2: Subject characteristics across UDBP:cR tertiles

```{r subchar udbpCr tertile}

subchar_table_tert <- tableone::CreateTableOne(
  vars = c("Age",
           "Sex",
           "Ethnicity",
           "BMI",
           "Waist",
           "eGFR",
           "ACR",
           "UrineCreatinine",
           "UrineMicroalbumin",
           "UrinaryCalcium",
           "UDBP",
           "udbpCr",
           "Systolic",
           "Diastolic",
           "MeanArtPressure",
           "dmStatus"),
  # strata = c("udbpCrTertile"),
  data = dsBase,
  factorVars = c("Sex", "Ethnicity", "dmStatus")
) %>% 
  print(nonnormal = c("UDBP",
                      "ACR",
                      "UrineMicroalbumin"),
        quote = FALSE,
        noSpaces = TRUE)
```

TABLE 3: Subject characteristics across time

```{r subchar_time}

subchar_table_time <- tableone::CreateTableOne(
  vars = c("Age",
           "Sex",
           "Ethnicity",
           "BMI",
           "Waist",
           "eGFR",
           "ACR",
           "UrineCreatinine",
           "UrineMicroalbumin",
           "UDBP",
           "udbpCr",
           "Systolic",
           "Diastolic",
           "MeanArtPressure",
           "dmStatus"),
  strata = c("fVN"),
  data = ds,
  factorVars = c("Sex", "Ethnicity", "dmStatus")
) %>% 
  print(nonnormal = c("UDBP",
                      "udbpCr",
                      "ACR",
                      "UrineMicroalbumin"),
        quote = FALSE,
        noSpaces = TRUE)
```

TABLE 4: Subject characteristics across time (complete data)

```{r subchar over time complete}

subchar_table_time <- tableone::CreateTableOne(
  vars = c("Age",
           "Sex",
           "Ethnicity",
           "BMI",
           "Waist",
           "eGFR",
           "ACR",
           "UrineCreatinine",
           "UrineMicroalbumin",
           "UDBP",
           "udbpCr",
           "Systolic",
           "Diastolic",
           "MeanArtPressure",
           "dmStatus"),
  strata = c("fVN"),
  data = dsComplete,
  factorVars = c("Sex", "Ethnicity", "dmStatus")
) %>% 
  print(nonnormal = c("UDBP",
                      "udbpCr",
                      "ACR",
                      "UrineMicroalbumin"),
        quote = FALSE,
        noSpaces = TRUE)
```

#### Incidence

```{r}

# desc_incidence() works for 0, 1 variables, not factors

ds %>% 
  dplyr::mutate(microalbumin = ifelse(acrStatus == "Microalbuminuria", 1, 0),
                macroalbumin = ifelse(acrStatus == "Macroalbuminuria", 1, 0)) %>% 
  dplyr::select(SID, VN, DM, microalbumin, macroalbumin) %>% 
  PROMISE.desc::desc_incidence("microalbumin")
```


### Cross-sectional

#### ACR

```{r acr boxplot}

# Clean data
acr <- ds %>% 
  dplyr::filter(fVN == "Baseline") %>% 
  dplyr::mutate(acrStatus2 = ifelse(acrStatus == "Normoalbuminuria", "Normoalbuminuria",
                                    "Albuminuria"),
                acrStatus2 = factor(acrStatus2,
                                    levels = c("Normoalbuminuria", "Albuminuria"),
                                    ordered = TRUE)) %>% 
  dplyr::select(acrStatus, acrStatus2, UDBP, udbpCr) %>%
  na.omit()

# Box plot of uVDBP in different albuminuria categories
acr %>% 
  box_plot_slides("acrStatus", "log(udbpCr)", 
            "Albuminuria",
            "log uVDBP:cr (μg/mmol)")

# n and values
acr %>% 
  group_by(acrStatus2) %>% 
  summarise(n = n(),
            median = median(log(udbpCr)),
            lower = quantile(log(udbpCr), probs=0.25),
            higher = quantile(log(udbpCr), probs=0.75))


give.n <- function(x){ return(c(ymin = median, label = length(x))) }
boxplot = boxplot + stat_summary(fun.data = give.n, geom = "text", fun.y = median)

# ANOVA
anova <- aov(formula = log(udbpCr)~acrStatus2, data = acr)
summary(anova)
TukeyHSD(anova)
rm(anova)
```


```{r acr scatterplot}

# Scatterplot of ACR and uVDBP ----------------------------------

dsBase %>% 
  scatter_plot_slides("log(ACR)", "log(udbpCr)",
               "log albumin:cr ratio (mg/mmol)",
               "log uVDBP:cr (μg/mmol)")

# Spearman Correlation ------------------------------------------

ds %>% 
  filter(fVN == "Baseline") %>% 
  cor.test(formula = ~ ACR + udbpCr, data = ., method = "spearman")

# Linear Regression ---------------------------------------------

dsBase %>% 
  prep_mason_data() %>% 
  mason_glm(y = "ACR",
            x = "udbpCr",
            covars = c("Age", "Sex", "Ethnicity", "dmStatus")
            ) %>% 
  gee_results_table() %>% 
  pander::pander()
  
```

#### eGFR

```{r eGFR boxplot}

# Clean data
eGFR <- ds %>% 
  filter(fVN == "Baseline") %>% 
  select(eGFRStatus, udbpCr, UDBP) %>%
  na.omit()

# Box plot of uVDBP in different eGFR categories
eGFR %>% 
  box_plot_slides("eGFRStatus", "log(udbpCr)", 
            "Kidney Status",
            "log uVDBP:cr (μg/mmol)")

# n and values
eGFR %>% 
  group_by(eGFRStatus) %>% 
  summarise(n = n(),
            median = median(log(udbpCr)),
            lower = quantile(log(udbpCr), probs=0.25),
            higher = quantile(log(udbpCr), probs=0.75))

# ANOVA
anova <- aov(formula = log(UDBP)~eGFRStatus, data = eGFR)
summary(anova)
TukeyHSD(anova)
rm(anova)
```

```{r eGFR scatterplot}

# Scatterplot of eGFR and uVDBP ----------------------------------

ds %>% 
  filter(fVN == "Baseline") %>% 
  scatter_plot_slides("log(eGFR)", "log(udbpCr)",
               "log eGFR (ml/min/1.73m^2)",
               "log uVDBP:cr (μg/mmol)")

# Spearman Correlation ------------------------------------------

ds %>% 
  filter(fVN == "Baseline") %>% 
  cor.test(formula = ~ eGFR + udbpCr, data = ., method = "spearman")

# Linear Regression ---------------------------------------------

dsBase %>% 
  prep_mason_data() %>% 
  mason_glm(y = "eGFR",
            x = "UDBP",
            covars = c("Age", "Sex", "Ethnicity", "dmStatus")
            ) %>% 
  gee_results_table(table = FALSE)
  
```

```{r}
ds %>% 
  prep_mason_data() %>% 
  mason_glm(y = "eGFR",
            x = "udbpCr"
            # covars = "ageBase"
            )
```

### uVDBP vs uVDBP:cr

```{r}

# Scatterplot of uVDBP and uVDBP:cr -----------------------------

ds %>% 
  scatter_plot("log(UDBP)", "log(udbpCr)",
               "log uVDBP",
               "log uVDBP:cr", facet = TRUE)

ds %>% 
  scatter_plot("log(UDBP)", "log(UrineCreatinine)",
               "log uVDBP",
               "log uCreatinine", facet = TRUE)

# Spearman Correlation ------------------------------------------

ds %>% 
  filter(fVN == "Baseline") %>% 
  cor.test(formula = ~ udbpCr + UDBP, data = ., method = "spearman")

```

```{r}

# Scatterplot of eGFR and uVDBP:cr -----------------------------

ds %>% 
  dplyr::filter(fVN == "Baseline") %>% 
  scatter_plot("log(eGFR)", "log(UDBP)",
               "log eGFR",
               "log uVDBP")

# Spearman Correlation ------------------------------------------

ds %>% 
  filter(fVN == "Baseline") %>% 
  cor.test(formula = ~ UDBP + eGFR, data = ., method = "spearman")

# Scatterplot of eGFR and uCr -----------------------------

ds %>% 
  dplyr::filter(fVN == "6Year") %>% 
  scatter_plot("log(eGFR)", "log(UrineCreatinine)",
               "log eGFR",
               "log creatinine")

# Spearman Correlation ------------------------------------------

ds %>% 
  filter(fVN == "6Year") %>% 
  cor.test(formula = ~ UrineCreatinine + eGFR, data = ., method = "spearman")

```

### Medication

* There are 705 values across all time points with blood pressure medication data
* This is approximately half of all observations (1852 vs 705)
* CHECK IF MISSINGNESS == NOT TAKING OR REALLY MISSING!!

```{r med}

ds_med <- ds %>% 
  dplyr::select(SID, VN, fMedsBP) %>% 
  na.omit()
```


### Progression

<!-- ACR -->

<!-- ```{r acr progression} -->

<!-- ds %>%  -->
<!--   plot_progress(yvar = "log(ACR)", -->
<!--                 ylab = "log(ACR)") -->

<!-- # Complete data -->

<!-- dsComplete %>%  -->
<!--   plot_progress(yvar = "log(ACR)", -->
<!--                 ylab = "log(ACR)") -->

<!-- # Boxplot ---------------------------------------------------------- -->

<!-- ds %>%  -->
<!--   plot_progress_boxplot(yvar = "log(ACR)") -->

<!-- # Complete data -->

<!-- dsComplete %>%  -->
<!--   plot_progress_boxplot(yvar = "log(ACR)", -->
<!--                         ylab = "log(ACR)") -->

<!-- # ANOVA ------------------------------------------------------------ -->

<!-- anova <- aov(formula = log(ACR)~fVN, data = ds) -->
<!-- summary(anova) -->
<!-- TukeyHSD(anova) -->
<!-- rm(anova) -->
<!-- ``` -->

<!-- eGFR -->

<!-- ```{r eGFR progression} -->

<!-- ds %>%  -->
<!--   plot_progress_boxplot(yvar = "log(eGFR)") -->

<!-- # Complete data -->

<!-- dsComplete %>%  -->
<!--   plot_progress_boxplot(yvar = "log(eGFR)") -->

<!-- # ANOVA -->

<!-- anova <- aov(formula = log(eGFR)~fVN, data = dsComplete) -->
<!-- summary(anova) -->
<!-- TukeyHSD(anova) -->
<!-- rm(anova) -->
<!-- ``` -->

<!-- UDBP -->

<!-- ```{r udbp progression} -->

<!-- ds %>%  -->
<!--   plot_progress_boxplot(yvar = "log(udbpCr)", -->
<!--                 ylab = "log(UDBP:cr)") -->

<!-- # Complete data -->

<!-- dsComplete %>%  -->
<!--   plot_progress_boxplot(yvar = "log(udbpCr)", -->
<!--                 ylab = "log(UDBP:cr)") -->

<!-- # ANOVA -->

<!-- anova <- aov(formula = log(udbpCr)~fVN, data = dsComplete) -->
<!-- summary(anova) -->
<!-- TukeyHSD(anova) -->
<!-- rm(anova) -->
<!-- ``` -->


<!-- ```{r udbp_time} -->

<!-- # Transformed using log -->
<!-- ds %>%  -->
<!--   dplyr::select(udbpCr, UDBP, fVN) %>%  -->
<!--   box_plot("fVN", "log(udbpCr)", -->
<!--            "Visit Number", "log UDBP:Creatinine") -->

<!-- # ANOVA Transformed -->
<!-- anova <- aov(formula = log(udbpCr)~fVN, data = ds) -->
<!-- summary(anova) -->
<!-- TukeyHSD(anova) -->
<!-- rm(anova) -->

<!-- # Untransformed -->
<!-- ds %>%  -->
<!--   dplyr::select(udbpCr, UDBP, fVN) %>%  -->
<!--   box_plot("fVN", "udbpCr", -->
<!--            "Visit Number", "UDBP:Creatinine") -->

<!-- # ANOVA Untransformed -->
<!-- anova <- aov(formula = udbpCr~fVN, data = ds) -->
<!-- summary(anova) -->
<!-- TukeyHSD(anova) -->
<!-- rm(anova) -->

<!-- # n -->
<!-- ds %>%  -->
<!--   group_by(fMedsBP) %>%  -->
<!--   summarise(n = n()) -->

<!-- # LOESS curve -->
<!-- ds %>%  -->
<!--   scatter_plot("VN", "log(UDBP)",  -->
<!--                "Visit Number", "log(UDBP)") -->

<!-- # Plot individual progression with mean -->
<!-- plot_progress_by(ds, byvar = "Sex") -->

<!-- # Plot subset of individual progression with mean -->
<!-- ds %>%  -->
<!--   dplyr::filter(Hypertension == 0) %>%  -->
<!--   plot_progress_data(byvar = "fMedsBP") -->

<!-- # Interaction between BP medication and hypertension -->
<!-- ds %>%  -->
<!--   prep_mason_data() %>%  -->
<!--   mason_gee(y = c("UDBP"), -->
<!--               x = "MedsBloodPressure", -->
<!--               covars = c("VN", "ageBase", "Sex", "Ethnicity", "BMI", "fDM", "Hypertension"), -->
<!--             intvar = "Hypertension") %>%  -->
<!--   dplyr::select(Yterms, Xterms, term, p.value, conf.low, conf.high, sample.total) -->
<!-- ``` -->

### Generalized Estimating Equations

```{r}
prep_mason_data_kidney <- ds %>%
    dplyr::mutate(
      udbpBase = ifelse(fVN == "Baseline", UDBP, NA),
      udbpCrBase = ifelse(fVN == "Baseline", udbpCr, NA),
      ageBase = ifelse(fVN == "Baseline", Age, NA),
      DM = ifelse(DM == 1, "DM", "notDM"),
      fDM = relevel(as.factor(DM), "notDM"),
      dmStatus = factor(dmStatus, ordered = FALSE),
      Ethnicity = ifelse(Ethnicity == "European", Ethnicity, "Other"),
      Ethnicity = relevel(as.factor(Ethnicity), "Other"),

      lACR = log(ACR),
      leGFR = log(eGFR)
    ) %>%

    dplyr::filter(!(fVN == "Baseline" &
                      acrStatus == "Macroalbuminuria")) %>%
    dplyr::filter(!(fVN == "Baseline" & eGFR < 60)) %>%
    dplyr::filter(!(fVN == "Baseline" & dmStatus == "DM")) %>%

    dplyr::mutate_each(dplyr::funs(as.numeric(scale(.))),
                       UDBP,
                       udbpBase,
                       udbpCr,
                       udbpCrBase,
                       MonthsFromBaseline,
                       ageBase) %>%

    dplyr::arrange(SID, fVN) %>%
    dplyr::group_by(SID) %>%
    tidyr::fill(udbpBase, udbpCrBase, ageBase) %>%
    dplyr::ungroup() %>%
    dplyr::arrange(SID, fVN)
```


GEE where predictor is baseline UDBP

```{r gee_baseline}

# Predictor is baseline UDBP ----------------------------------------

gee_kidney_baseline1 <- ds %>% 
  prep_mason_data_kidney() %>% 
  mason_geeplot(y = c("lACR", "leGFR"),
            x = "udbpCrBase",
            covars = c("YearsFromBaseline", "ageBase", "Ethnicity", 
                       "dmStatus")) %>%
  mason::polish_renaming(rename_gee_kidney)

gee_kidney_baseline2 <- ds %>% 
  prep_mason_data_kidney() %>% 
  mason_geeplot(y = c("lACR", "leGFR"),
            x = "udbpCrBase",
            covars = c("YearsFromBaseline", "ageBase", "Sex", "Ethnicity", 
                       "dmStatus")) %>%
  mason::polish_renaming(rename_gee_kidney)

# GEE table ----------------------------------------------------------

gee_kidney_baseline2 %>% 
  gee_results_table() %>% 
  pander::pander()

# Plot ---------------------------------------------------------------

plot_gee_results_kidney_base(gee_kidney_baseline2,
                 yvars = c("ACR (mg/mmol)", 
                           "eGFR (ml/min/1.73m^2)"))
```

GEE where predictor is UDBP over time

```{r gee}
# Predictor is UDBP at all time points ------------------------------

gee_kidney1 <- ds %>% 
  prep_mason_data_kidney() %>% 
  mason_geeplot(y = c("lACR", "leGFR"),
            x = "udbpCr",
            covars = c("YearsFromBaseline", "ageBase", "Ethnicity", 
                       "dmStatus")) %>%
  mason::polish_renaming(rename_gee_kidney)

gee_kidney2 <- ds %>% 
  prep_mason_data_kidney() %>% 
  mason_geeplot(y = c("lACR", "leGFR"),
            x = "udbpCr",
            covars = c("YearsFromBaseline", "ageBase", "Sex", "Ethnicity", 
                       "dmStatus")) %>%
  mason::polish_renaming(rename_gee_kidney)

# GEE table ----------------------------------------------------------

gee_kidney2 %>% 
  gee_results_table() %>% 
  pander::pander()

# Plot ---------------------------------------------------------------

plot_gee_results_kidney(gee_kidney2,
                 yvars = c("ACR (mg/mmol)", 
                           "eGFR (ml/min/1.73m^2)"))
```


<!-- ### Interaction with Time -->

<!-- ```{r interaction_vn} -->

<!-- # Unadjusted interaction with visit number -->
<!-- ds %>%  -->
<!--   prep_mason_data() %>%  -->
<!--   mason_gee(y = c("ACR", "eGFR"), -->
<!--               x = "UDBP", -->
<!--               covars = c("VN"), -->
<!--               intvar = "VN") %>%  -->
<!--   dplyr::select(Yterms, Xterms, term, p.value, conf.low, conf.high, sample.total) -->

<!-- # Adjusted interaction with visit number -->
<!-- ds %>%  -->
<!--   prep_gee_data() %>%  -->
<!--   analyze_gee(y = c("ACR", "eGFR"), -->
<!--               x = "UDBP", -->
<!--               covars = c("VN", "ageBase", "Sex", "Ethnicity", "BMI", "fDM"), -->
<!--               intvar = "VN") %>%  -->
<!--   dplyr::select(Yterms, Xterms, term, p.value, conf.low, conf.high, sample.total) -->

<!-- # Unadjusted interaction with months from baseline -->
<!-- ds %>%  -->
<!--   prep_mason_data() %>%  -->
<!--   mason_gee(y = c("ACR", "eGFR"), -->
<!--               x = "UDBP", -->
<!--               covars = c("MonthsFromBaseline"), -->
<!--               intvar = "MonthsFromBaseline") %>%  -->
<!--   dplyr::select(Yterms, Xterms, term, p.value, conf.low, conf.high, sample.total) -->

<!-- # Adjusted interaction with months from baseline -->
<!-- ds %>%  -->
<!--   prep_mason_data() %>%  -->
<!--   mason_gee(y = c("ACR", "eGFR"), -->
<!--               x = "UDBP", -->
<!--               covars = c("MonthsFromBaseline", "ageBase", "Sex", "Ethnicity", "BMI", "fDM"), -->
<!--               intvar = "MonthsFromBaseline") %>%  -->
<!--   dplyr::select(Yterms, Xterms, term, p.value, conf.low, conf.high, sample.total) -->
<!-- ``` -->
