---
output:
  word_document: default
  html_document: default
---

# Title page #

Title: Urinary Vitamin D Binding Protein Loss as a Potential Biomarker for Diabetic Nephropathy in Patients at Risk for Type 2 Diabetes

Author: Windy ZN Wang

Affiliations: University of Toronto

Disclaimers:

Correspondence: windy.wang@mail.utoronto.ca

Funding support: CIHR, Sharon Zeiler, BBDC Novo Nordisk, Dairy Farmers

Abstract
========

Background
==========

Methods
=======

Subjects
--------

<!--
## Load data ##
-->

```{r setup, collapse=TRUE, echo=FALSE, message=FALSE, warning=FALSE}

# Only set if the Rmd file is not in the parent directory (ie. 'projectname/')
knitr::opts_knit$set(root.dir = '../')

knitr::opts_chunk$set(collapse = TRUE, echo = FALSE, message = FALSE, warning = FALSE)
library(dplyr)
devtools::load_all()
load_data(update = TRUE)
set_options()
# extrafont::loadfonts(device="win")
```

```{r loadData, message=FALSE}

# source('.Rprofile')
# run_setup()
# load_data('data/project_data.rda')
# dim(ds)
ds <- project_data

dsBase <- ds %>% 
  dplyr::filter(fVN == "Baseline")

# dsComplete <- ds %>% 
#   dplyr::group_by(SID) %>%
#   dplyr::filter(n() == 3) %>%
#   dplyr::ungroup()
```

```{r caption_setup}

## Include captions below using `captioner` package

figNums <- captioner::captioner(prefix = 'Figure')
supFigNums <- captioner::captioner(prefix = 'Supplementary Figure')
tabNums <- captioner::captioner(prefix = 'Table')
supTabNums <- captioner::captioner(prefix = 'Supplementary Table')
```

PAPER 1: KIDNEY MEASURE RESULTS
===============================

### Subject Characteristics

TABLE 1: Subject characteristics across UDBP tertiles

```{r subchar tertile}

subchar_table_tert <- tableone::CreateTableOne(
  vars = c("Age",
           "Sex",
           "Ethnicity",
           "BMI",
           "Waist",
           "eGFR",
           "ACR",
           "UrineCreatinine",
           "UrineMicroalbumin",
           "UrinaryCalcium",
           "UDBP",
           "Systolic",
           "Diastolic",
           "MeanArtPressure",
           "dmStatus"),
  strata = c("UDBPTertile"),
  data = dsBase,
  factorVars = c("Sex", "Ethnicity", "dmStatus")
) %>% 
  print(nonnormal = c("UDBP",
                      "ACR",
                      "UrineMicroalbumin"),
        quote = FALSE,
        noSpaces = TRUE)
```

```{r subchar over time}

subchar_table_time <- tableone::CreateTableOne(
  vars = c("Age",
           "Sex",
           "Ethnicity",
           "BMI",
           "Waist",
           "eGFR",
           "ACR",
           "UrineCreatinine",
           "UrineMicroalbumin",
           "UrinaryCalcium",
           "UDBP",
           "Systolic",
           "Diastolic",
           "MeanArtPressure",
           "dmStatus"),
  strata = c("fVN"),
  data = ds,
  factorVars = c("Sex", "Ethnicity", "dmStatus")
) %>% 
  print(nonnormal = c("UDBP",
                      "ACR",
                      "UrineMicroalbumin"),
        quote = FALSE,
        noSpaces = TRUE)
```


Paper 1: Kidney Measures
------------------------

### Cross-sectional

Albuminuria as assessed by albumin-to-creatinine ratio

```{r acr}

# Clean data
acr <- ds %>% 
  filter(fVN == "Baseline") %>% 
  select(acrStatus, udbpCrRatio) %>%
  na.omit()

# Box plot of uVDBP in different albuminuria categories
acr %>% 
  box_plot("acrStatus", "log(udbpCrRatio)", 
            "Albuminuria",
            "log uVDBP:Creatinine")

# n
acr %>% 
  group_by(acrStatus) %>% 
  summarise(n = n())

# ANOVA
anova <- aov(formula = log(udbpCrRatio)~acrStatus, data = acr)
summary(anova)
TukeyHSD(anova)
rm(anova)
```


```{r acr scatterplot}

# Scatterplot of ACR and uVDBP ----------------------------------

ds %>% 
  filter(fVN == "Baseline") %>% 
  scatter_plot("log(ACR)", "log(udbpCrRatio)",
               "log Albumin:Creatinine Ratio",
               "log UDBP:Creatinine")

# Spearman Correlation ------------------------------------------

ds %>% 
  filter(fVN == "Baseline") %>% 
  cor.test(formula = ~ ACR + udbpCrRatio, data = ., method = "spearman")

# Linear Regression ---------------------------------------------

ds %>% 
  prep_mason_data() %>% 
  mason_glm(y = "ACR",
            x = "udbpCrRatio",
            covars = c("ageBase", "Sex", "Ethnicity", "fDM", "fMedsBP")
            )
  
```

eGFR

```{r eGFR}

# Clean data
eGFR <- ds %>% 
  filter(fVN == "Baseline") %>% 
  select(eGFRStatus, udbpCrRatio) %>%
  na.omit()

# Box plot of uVDBP in different eGFR categories
eGFR %>% 
  box_plot("eGFRStatus", "log(udbpCrRatio)", 
            "Kidney Status",
            "log uVDBP:Creatinine")

# n
eGFR %>% 
  group_by(eGFRStatus) %>% 
  summarise(n = n())

# ANOVA
anova <- aov(formula = log(udbpCrRatio)~eGFRStatus, data = eGFR)
summary(anova)
TukeyHSD(anova)
rm(anova)
```

```{r eGFR scatterplot}

# Scatterplot of eGFR and uVDBP ----------------------------------

ds %>% 
  filter(fVN == "Baseline") %>% 
  scatter_plot("log(eGFR)", "log(udbpCrRatio)",
               "log Estimated Glomerular Filtration Rate",
               "log UDBP:Creatinine")

# Spearman Correlation ------------------------------------------

ds %>% 
  filter(fVN == "Baseline") %>% 
  cor.test(formula = ~ eGFR + udbpCrRatio, data = ., method = "spearman")

# Linear Regression ---------------------------------------------

ds %>% 
  prep_mason_data() %>% 
  mason_glm(y = "eGFR",
            x = "udbpCrRatio",
            covars = c("ageBase", "Sex", "Ethnicity", "fDM", "fMedsBP")
            )
  
```

```{r}
ds %>% 
  prep_mason_data() %>% 
  mason_glm(y = "eGFR",
            x = "udbpCrRatio"
            # covars = "ageBase"
            )
```


Medication

* There are 705 values across all time points with blood pressure medication data
* This is approximately half of all observations (1852 vs 705)
* CHECK IF MISSINGNESS == NOT TAKING OR REALLY MISSING!!

```{r med}

ds_med <- ds %>% 
  dplyr::select(SID, VN, fMedsBP) %>% 
  na.omit()
```


### Progression

ACR

```{r acr progression}

ds %>% 
  plot_progress(yvar = "log(ACR)",
                ylab = "log(ACR)")
```

eGFR
```{r eGFR progression}

ds %>% 
  plot_progress(yvar = "log(eGFR)",
                ylab = "log(eGFR)")
```

UDBP
```{r udbp progression}

ds %>% 
  plot_progress(yvar = "log(UDBP)",
                ylab = "log(UDBP)")
```


```{r udbp_time}

# Transformed using log
ds %>% 
  dplyr::select(udbpCrRatio, UDBP, fVN) %>% 
  box_plot("fVN", "log(udbpCrRatio)",
           "Visit Number", "log UDBP:Creatinine")

# ANOVA Transformed
anova <- aov(formula = log(udbpCrRatio)~fVN, data = ds)
summary(anova)
TukeyHSD(anova)
rm(anova)

# Untransformed
ds %>% 
  dplyr::select(udbpCrRatio, UDBP, fVN) %>% 
  box_plot("fVN", "udbpCrRatio",
           "Visit Number", "UDBP:Creatinine")

# ANOVA Untransformed
anova <- aov(formula = udbpCrRatio~fVN, data = ds)
summary(anova)
TukeyHSD(anova)
rm(anova)

# n
ds %>% 
  group_by(fMedsBP) %>% 
  summarise(n = n())

# By variable
ds %>%
  # dplyr::filter(MedsBloodPressure == 0) %>%
  line_plot("fVN", "log(UDBP)", "Sex", "Visit Number", "log(UDBP)")

# LOESS curve
ds %>% 
  scatter_plot("VN", "log(UDBP)", 
               "Visit Number", "log(UDBP)")

# Plot individual progression with mean
plot_progress_data_by(ds, byvar = "Hypertension")

# Plot subset of individual progression with mean
ds %>% 
  dplyr::filter(Hypertension == 0) %>% 
  plot_progress_data(byvar = "fMedsBP")

# Interaction between BP medication and hypertension
ds %>% 
  prep_mason_data() %>% 
  mason_gee(y = c("UDBP"),
              x = "MedsBloodPressure",
              covars = c("VN", "ageBase", "Sex", "Ethnicity", "BMI", "fDM", "Hypertension"),
            intvar = "Hypertension") %>% 
  dplyr::select(Yterms, Xterms, term, p.value, conf.low, conf.high, sample.total)
```

### Prospective Analysis

Generalized Estimating Equations

```{r gee_analysis}

# Predictor is baseline UDBP ----------------------------------------

gee <- ds %>% 
  prep_mason_data() %>% 
  mason_gee(y = c("ACR", "eGFR"),
            x = "udbpBase",
            covars = c("VN", "ageBase", "Sex", "Ethnicity", "BMI", "fDM")) 
# %>% 
#   dplyr::filter(!term == "(Intercept)") %>% 
#   dplyr::select(Yterms, Xterms, term, estimate, p.value, conf.low, conf.high, std.error)

# Predictor is UDBP at all time points ------------------------------

gee <- ds %>% 
  prep_mason_data() %>% 
  mason_gee(y = c("ACR", "eGFR"),
              x = "UDBP",
              covars = c("VN", "ageBase", "Sex", "Ethnicity", "BMI", "fDM")) 
# %>% 
#   dplyr::filter(!term == "(Intercept)") %>% 
#     dplyr::select(Yterms, Xterms, term, estimate, p.value, conf.low, conf.high, std.error)
```

```{r gee_plot}

plot_gee_results(gee,
                 yvars = c("ACR", "eGFR"))
```


### Interaction with Time

```{r interaction_vn}

# Unadjusted interaction with visit number
ds %>% 
  prep_mason_data() %>% 
  mason_gee(y = c("ACR", "eGFR"),
              x = "UDBP",
              covars = c("VN"),
              intvar = "VN") %>% 
  dplyr::select(Yterms, Xterms, term, p.value, conf.low, conf.high, sample.total)

# Adjusted interaction with visit number
ds %>% 
  prep_gee_data() %>% 
  analyze_gee(y = c("ACR", "eGFR"),
              x = "UDBP",
              covars = c("VN", "ageBase", "Sex", "Ethnicity", "BMI", "fDM"),
              intvar = "VN") %>% 
  dplyr::select(Yterms, Xterms, term, p.value, conf.low, conf.high, sample.total)

# Unadjusted interaction with months from baseline
ds %>% 
  prep_mason_data() %>% 
  mason_gee(y = c("ACR", "eGFR"),
              x = "UDBP",
              covars = c("MonthsFromBaseline"),
              intvar = "MonthsFromBaseline") %>% 
  dplyr::select(Yterms, Xterms, term, p.value, conf.low, conf.high, sample.total)

# Adjusted interaction with months from baseline
ds %>% 
  prep_mason_data() %>% 
  mason_gee(y = c("ACR", "eGFR"),
              x = "UDBP",
              covars = c("MonthsFromBaseline", "ageBase", "Sex", "Ethnicity", "BMI", "fDM"),
              intvar = "MonthsFromBaseline") %>% 
  dplyr::select(Yterms, Xterms, term, p.value, conf.low, conf.high, sample.total)
```



PAPER 2: VITAMIN D RESULTS
==========================

### Subject Characteristics

```{r subchar vitamind baseline}

# Diet information is not available at baseline. Only VN 3 data is available at the present time.

subchar_vitd_base <- tableone::CreateTableOne(
  vars = c("Age",
           "Sex",
           "Ethnicity",
           "BMI",
           "Waist",
           "eGFR",
           "ACR",
           "UrineCreatinine",
           "UrineMicroalbumin",
           "UrinaryCalcium",
           "UDBP",
           "VitaminD",
           "PTH",
           "Systolic",
           "Diastolic",
           "MeanArtPressure",
           "diet_milk",
           "diet_cal",
           "diet_supp_cal",
           "diet_supp_vitd",
           "OralContraceptive",
           "dmStatus"),
  strata = c("vitdStatus"),
  data = dsBase,
  factorVars = c("Sex", "Ethnicity", "OralContraceptive", "dmStatus")
) %>% 
  print(nonnormal = c("UDBP",
                      "ACR",
                      "UrineMicroalbumin"),
        quote = FALSE,
        noSpaces = TRUE)
```


# Discussion #

## Acknowledgements ##

# References #
